Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down

### 2022-08-24 paper/ms.tex
Updates to text

\section{Introduction}
Approximately two-thirds of spiral disks host an elongated bar-like feature at
their center \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our own
Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of bars
is not difficult to explain, since stellar disks simulated in isolation almost
always form bar-like structures \citep{1971ApJ...168..343H}. Several studies
have shown that a spherical potential, such as a stellar bulge or dark matter
halo, acts to stabilize a stellar disk against the bar instability
\citep{1973ApJ...186..467O, 1976AJ.....81...30H}.

@@

\begin{figure*}
    \centering
    \includegraphics[width=18cm]{fig/mockHST.pdf}
    \caption{Synthetic Hubble Space Telescope images of our simulations with and
    without the interstellar medium. The upper panels show an \Nbody{} only
    simulation while the lower panels show a simulation which includes the
    SMUGGLE model for the interstellar medium. Each panel is $20\,\textrm{kpc}$
    to a side. Columns show different points in time, separated by
    $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
    length and strength. In the SMUGGLE run, the bar remains at approximately
    the same length and strength over the course of the simulation. The \Nbody{}
    model is identical to the GALAKOS model, discussed in the text. Details on
    the production of these synthetic images is given in Appendix~\ref{app:hst}.}\label{fig:overview}
\end{figure*}

@@

\section{Comparison to the Milky Way}
\label{app:milkyway}
Our fiducial disk model is a relatively good  model of the Milky Way for several
Gyr. This is uncommon in models of galaxies that include the gas phase of the
disk but no circumgalactic medium. As mentioned earlier, the pattern speed seems
to match the observed pattern speed of the Milky Way's bar
\citep{2019MNRAS.490.4740B}. We briefly summarize some of the other ways our disk is comparable to the Milky Way.

We computed the circular velocity curve of our model using the \texttt{AGAMA}
package\cite{2019MNRAS.482.1525V}. We fit the baryonic component (stellar
disk, bulge, gas, and newly formed stars) with an axisymmetric cylindrical
spline with $20$ grid points in both the radial and vertical direction
spanning $0.2$ to $50\,\textrm{kpc}$ in the radial direction and from $0.02$
to $10\,\textrm{kpc}$ in the vertical direction. We fit the dark matter halo
using a spherically symmetric multipole fit with a maximum angular harmonic
coefficient of $l=2$. We plot the circular velocity curve in
Fig.~\ref{fig:vcirc} compared to observational
estimates\cite{2019ApJ...871..120E}. The SMUGGLE disk (which includes
additional mass in the form of gas) has a slightly higher circular velocity
than the \Nbody{} disk which, itself, is slightly higher than the
observational estimates.

>>>>

\section{Introduction}
\the\columnwidth

\the\textwidth

% \showthe\textwidth
Approximately two-thirds of spiral disks host an elongated bar-like feature at
their center \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our own
Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of bars
is not difficult to explain, since stellar disks simulated in isolation almost
always form bar-like structures \citep{1971ApJ...168..343H}. Several studies
have shown that a spherical potential, such as a stellar bulge or dark matter
halo, acts to stabilize a stellar disk against the bar instability
\citep{1973ApJ...186..467O, 1976AJ.....81...30H}.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fig/mockHST.pdf}
    \caption{Synthetic Hubble Space Telescope images of our simulations with and
    without the interstellar medium. The upper panels show an \Nbody{} only
    simulation while the lower panels show a simulation which includes the
    SMUGGLE model for the interstellar medium. Each panel is $20\,\textrm{kpc}$
    to a side. Columns show different points in time, separated by
    $1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
    length and strength. In the SMUGGLE run, the bar remains at approximately
    the same length and strength over the course of the simulation. The \Nbody{}
    model is identical to the GALAKOS model, discussed in the text. Details on
    the production of these synthetic images is given in Appendix~\ref{app:hst}.}\label{fig:overview}
\end{figure*}

@@

\section{Comparison to the Milky Way}
\label{app:milkyway}
For several Gyr, our fiducial disk exhibits several properties in reasonable
agreement to the Milky Way. This is uncommon in models of galaxies that include
the gas phase of the disk but no circumgalactic medium. As mentioned earlier,
the pattern speed seems to match the observed pattern speed of the Milky Way's
bar \citep{2019MNRAS.490.4740B}. We briefly summarize some of the other ways our
disk is comparable to the Milky Way.

We computed the circular velocity curve of our model using the \texttt{AGAMA}
package\cite{2019MNRAS.482.1525V}. We fit the baryonic component (stellar disk,
bulge, gas, and newly formed stars) with an axisymmetric cylindrical spline with
$20$ grid points in both the radial and vertical direction spanning $0.2$ to
$50\,\textrm{kpc}$ in the radial direction and from $0.02$ to $10\,\textrm{kpc}$
in the vertical direction. We fit the dark matter halo using a spherically
symmetric multipole fit with a maximum angular harmonic coefficient of $l=2$. We
plot the circular velocity curve in Fig.~\ref{fig:vcirc} compared to
observational estimates\cite{2019ApJ...871..120E}. The SMUGGLE disk (which
includes additional mass in the form of gas) has a slightly higher circular
velocity than the \Nbody{} disk which, itself, is slightly higher than the
observational estimates. Overall, though, the circular velocity curves between
our model and that observed in the Milky Way are broadly consistent.
